\documentclass{report}
\usepackage{geometry}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{multicol}
\usepackage{blindtext}
\usepackage{color}
\usepackage[fontsize=16pt]{fontsize}
\usepackage{lipsum}

\setlength{\columnsep}{1cm}
\def\columnseprulecolor{\color{blue}}
\date{Fall 2023}

\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}


\title{PHYS110}
\author{Aiden Sirotkine}

\begin{document}

\pagestyle{fancy}
\maketitle
\tableofcontents
\clearpage

\chapter{Overview}
\begin{itemize}
\item
Only an 8 week pass/fail course

\item
In-Class

Lectures, Guest Speakers, Student/Faculty Panels

\item
Out of Class

Peer Mentor Meets, Safety Training. 
\end{itemize}

\section*{Necessary Things}
\begin{itemize}
\item
Attend and participate in 5 of 8 classes

\item
complete a 4 year academic plan

\item
Do Research and Lab Safety Training

\end{itemize}


\section*{attendance}
\begin{itemize}
\item
Confirm attendance using Microsoft Forms QR Code

\item
Can send in absence form, but only 12 days in advance

\item
If you are sick, DON'T COME TO CLASS

Attend via Zoom and email professor

\end{itemize}

\section*{Peer and Mentor Programs}
\begin{itemize}
\item
Faculty Mentors will tell you what the good classes and resources are. Internships, study groups, opportunities, etc.

\item
Peer mentor will look at your 4-year plan and tell you if you are crazy or not. 

\end{itemize}

\newpage 
\section*{Physics Student Organizations}
\begin{itemize}
\item
Society of Physics Students:

Social, Academic, and Professional Club. Weekly Goofy Physics Meetups. Engineering Open House. Trip to Argonne National Labs????? 

\item
Society for Women in Physics

\item 
Physics Student Advisory Board (PhySAB)

Undergrad Lounge, Communication between students and departments, Research Symposium, Study Hall Events(?)

\item
 Society for Underrepresented Physics Students, 
 
 buncha people who just vibe
 
 \item
 PhsyVan
 
 Magic Show

\item 
Guidance for Physics Students. 

\end{itemize}


\end{document}